---
title: "Time Series HW 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Chapter 1 Exercises

```{r}
library(TSA)
```

# 1.1

printing it to see how it looks

```{r}
data(larain)
print(larain)
```

looks like it's a data type called Time Series, from the TSA library. checking its type.

```{r}
typeof(larain)
```

Why?

```{r}
plot(larain)
```

plotting the way the book does it.

```{r}
# setting the frame as square but it doesn't work
# win.graph(width=3,height=3,pointsize=8)
```

```{r}
plot(y = larain, x=zlag(larain), ylab='Inches of rain', xlab='Prev. Year Inches of rain')
```

# 1.2

```{r}
data(color)
color
```

```{r}
typeof(color)
```

```{r}
plot(color,ylab='Color Property',xlab='Batch',type='o')
```

# 1.3

iid sample of the Gaussian dist

```{r}
ex = rnorm(48)
print(ex)
typeof(ex)
```

```{r}
plot(ex, ylab = "Value", xlab= "Index", type = "o")
```

```{r}
#dev.off() #the guys at the company told me https://support.posit.co/hc/en-us/articles/200488548-Problem-with-Plots-or-Graphics-Device-in-the-RStudio-IDE
# https://stackoverflow.com/questions/23050928/error-in-plot-new-figure-margins-too-large-scatter-plot
par(mfrow = c(5,3), mar=c(1.5,1.5,1.5,1.5))
for (i in 1:15){
  plot(rnorm(48),main=paste('Experiment #', i), xlab = 'Index', ylab = 'Value', type = 'o')
}
```

they all look like mere noise.

# 1.4

iid sample of Chi-squared dist

```{r}
plot(rchisq(48, 2), main="48 random chi2 (df=2) events", xlab='event', ylab='value', type='o')
```

```{r}
par(mfrow = c(5,3), mar=c(1.5,1.5,1.5,1.5))
for (i in 1:15){
  plot(rchisq(48,2),main=paste('Experiment #', i), xlab = 'Index', ylab = 'Value', type = 'o')
}
```

```{r, fig.height=10}
par(mfrow = c(5,3))
for (i in 1:15){
  y = rchisq(48,2)
  qqnorm(y, main=paste('Experiment #', i), xlab = 'Index', ylab = 'Value')
  qqline(y)
}
```

they all look random but not normal.

# 1.5

for a iid sample of t-dist

```{r}
plot(rt(48,5), main="48 iid t-Student events", xlab='event', ylab='value', type='o')
```

Drawing the differential plot next to time series plot to see if there is any temporal connection for the t-student dist.

```{r}
par(mfrow = c(5,2), mar=c(1.5,1.5,1.5,1.5))
for (i in 1:5){
  ex = rt(48,5)
  plot(ex,main=paste('Experiment #', i), xlab = 'Index', ylab = 'Value', type = 'o')
  plot(y=ex, x=zlag(ex), main='againt prev', ylab = "Value", xlab= "Previous")
}
```

```{r, fig.height=10}
par(mfrow = c(5,3))
for (i in 1:15){
  ex = rt(48,5)
  qqnorm(ex, main=paste('Experiment #', i), xlab = 'Index', ylab = 'Value')
  qqline(ex)
}
```

moving around zero with some outliers.

# 1.6

```{r}
data(tempdub)
print(tempdub)
typeof(tempdub)
```

```{r}
plot(tempdub,main='Mountly Temp of Dubuque', ylab='Temprature', type='o')
points(y=tempdub,x=time(tempdub), pch=as.vector(season(tempdub)))
```

Data expresses seasonality.
