---
title: "Time Series Chapter 7 HW"
author: "Pooria Assarehha"
date: "2024-01-06"
output: pdf_document
---

## Ex 7.1

-- $n = 100$
-- $r_1 = 0.8$
-- $r_2 = 0.5$
-- $r_3 = 0.4$
-- $\bar{Y} = 2$
-- $s^2 = 5$
-- $Y_t \sim$ AR(2) + $\mu$

-- $\hat{\phi}_1, \hat{\phi}_2, \hat{\theta}_0, \hat{\sigma}^2_e = ?$

$\hat{\theta}_0 = \bar{Y} = 2$

Using Yule-Walker Estimations:

$$
\begin{cases}
r_1 = \phi_1 + r_2\phi_2 \\
r_2 = r_1\phi_1 + \phi_2 
\end{cases} \quad \Rightarrow \quad \hat{\phi}_1 = \frac{r_1(1-r_2)}{1-r_2^2}  = 0.5\bar{3}  \quad , \quad \hat{\phi}_2 = \frac{r_2-r_1^2}{1-r_1^2} \approx -0.39 
$$

$$\hat{\sigma}_e^2 = \left(1 - \hat{\phi}_1r_1 - r_2\hat{\phi}_2\right)s^2 = 3.838$$

## Ex 7.3
-- $Y_t \sim$ AR(1)
-- $\phi \approx 0.7$
-- $Z_{0.975} \hat{Var}(\phi) \leq  0.1 \Rightarrow  n=?$

$$
AR(1) \Rightarrow \hat{\phi} = r \Rightarrow Var(r_1) = \frac{1-\phi_1^2}{n} \quad \Rightarrow 1.96\times \frac{1-0.7^2}{n} \leq 0.1  \Rightarrow n \geq \lceil 19.6 \times 0.51\rceil \Rightarrow n\geq 10
$$

## Ex 7.4

-- $Y_t \sim$ MA(1)
-- $\mu = 0$
-- $n=3$
-- $Y_1 = 0$
-- $Y_2 = -1$
-- $Y_3 = \frac{1}{2}$
-- (a) Cond-LS $\hat{\theta} = \frac{1}{2}$
-- (b) $\hat{\sigma}^2_e = ?$

$$e_t = Y_t + \theta e_{t-1} \begin{cases} e_1 = Y_1 + 0 \\ e_2 = Y_2 + \theta e_1 \\ e_3 = Y_3 + \theta e_2 \end{cases}$$

## Ex 7.6

Model I: $Y_t - \mu = \phi(Y_{t-1}-\mu) + e_t$
Model II: $Y_t = \phi Y_{t-1}+ \theta_0 + e_t $



## Ex 7.10

for a simulated MA(1) with $\theta = -0.6$
-- Method of Moments: $\hat{\theta} = ?$
-- Con LS of the prior and compare? 
-- MLE of the prior and compare?

we have to negate $\theta$ due to difference in MA definition 

```{r}
library(TSA)
yt = arima.sim(list(ma=c(0.6)), n=36)
plot(yt)
```



```{r}
r = acf(yt, ci.type='ma', plot = FALSE)
rho = ARMAacf(ma = c(0.6))
```


