---
title: "Time Series Chapter 8 HW"
author: "Pooria Assarehha"
date: "2024-01-20"
output: pdf_document
---

## Ex 8.1

$$Var(\hat{r}_k) = \frac{\phi^2}{n} \Rightarrow 0 \notin (0.5 - 1.96\times \sqrt{\frac{0.5^2}{100}}) = (0.5 \pm 0.1)$$
we must reject $\hat{r}_1 = 0$ meaning the residuals of the model are correlated. 

we want the residuals to be small and as independent as they can be.

## Ex 8.2

This is the same as 8.1 because for an MA(1) model $Var(\hat{r}_k) = \frac{\theta^2}{n}$

## Ex 8.4

```{r, message=FALSE}
library(TSA)
yt = arima.sim(list(ar=0.5), n = 30)
ar1 = arima(yt, order=c(1,0,0))
stdres = rstandard(ar1)
plot(stdres)
abline(h=0)
```
No info can be gained regarding normality. non normality cannot be rejected.

```{r}
qqnorm(stdres)
qqline(stdres)
```

We can Reject normality.

```{r}
acf(stdres)
```

```{r}
LB.test(ar1, lag=8)
```

## Ex 8.8

```{r}
data(oilfilters)
oil =arima(oilfilters,order=c(1,0,0))
oil
```
Yes it is. 

```{r}
acf(as.numeric(rstandard(oil)))
```

lag 12 has significant autocorrelation. this model did not consider seasonality.
